---
title: "String Manipulation with stringr in R"
author: "Hernan Perci Nuñez Palomino"
date: "26/2/2021"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float:
      collapsed: false
      smooth_scroll: true
        
---

La informacion que se presenta `pertenece` al curso descrito en el titulo de la plataforma **Datacamp**.

El archivo se encuentra alojado y actualizado en la pagina web <https://github.com/HernanPerci/String-Manipulation-with-stringr-in-R>

# Librerias usadas

```{r}

library(tidyverse)
library(babynames)
library(ggthemes)
library(rebus)
```


# Conceptos basicos de cadenas

## Comillas

```{r}

# Define line1
line1 <- "The table was a large one, but the three were all crowded together at one corner of it:"

# Define line2
line2 <- '"No room! No room!" they cried out when they saw Alice coming.'

# Define line3
line3 <- "\"There's plenty of room!\" said Alice indignantly, and she sat down in a large arm-chair at one end of the table."

```


## Print vs writelines
<http://www.unicode.org/charts/>

```{r}

# Putting lines in a vector
lines <- c(line1, line2, line3)

# Print lines
lines

# Use writeLines() on lines
writeLines(lines)

# Write lines with a space separator
writeLines(lines, sep = " ")

# Use writeLines() on the string "hello\n\U1F30D"
writeLines("hello\n\U1F30D")

```


## Sequencias de escape

```{r}

# Should display: To have a \ you need \\
writeLines("To have a \\ you need \\\\")

# Should display: 
# This is a really 
# really really 
# long string
writeLines("This is a really\nreally really\nlong string")

# Use writeLines() with 
# "\u0928\u092e\u0938\u094d\u0924\u0947 \u0926\u0941\u0928\u093f\u092f\u093e"
writeLines("\u0928\u092e\u0938\u094d\u0924\u0947 \u0926\u0941\u0928\u093f\u092f\u093e")

```


## Usando format() con numeros para convertirlos en cadena

```{r}

# Some vectors of numbers
percent_change  <- c(4, -1.91, 3.00, -5.002)
income <-  c(72.19, 1030.18, 10291.93, 1189192.18)
p_values <- c(0.12, 0.98, 0.0000191, 0.00000000002)

# Format c(0.0011, 0.011, 1) with digits = 1
format(c(0.0011, 0.011, 1), digits = 1)

# Format c(1.0011, 2.011, 1) with digits = 1
format(c(1.0011, 2.011, 1), digits = 1)

# Format percent_change to one place after the decimal point
format(percent_change, digits = 2)

# Format income to whole numbers
format(income, digits = 2)

# Format p_values in fixed format
format(p_values, scientific = FALSE)

```


## Controlando otros aspectos de la cadena

```{r}

formatted_income <- format(income, digits = 2)

# Print formatted_income
formatted_income

# Call writeLines() on the formatted income
writeLines(formatted_income)

# Define trimmed_income
trimmed_income <- format(income, digits = 2, trim = TRUE)

# Call writeLines() on the trimmed_income
writeLines(trimmed_income)

# Define pretty_income
pretty_income <- format(income, digits = 2, big.mark = ",")

# Call writeLines() on the pretty_income
writeLines(pretty_income)

```


## formatC

```{r}

# From the format() exercise
x <- c(0.0011, 0.011, 1)
y <- c(1.0011, 2.011, 1)

# formatC() on x with format = "f", digits = 1
formatC(x, format = "f", digits = 1)

# formatC() on y with format = "f", digits = 1
formatC(y,format = "f", digits = 1)

# Format percent_change to one place after the decimal point
format(percent_change, digits = 2)

# percent_change with flag = "+"
pretty_percent <- formatC(percent_change, format = "f", digits = 1, flag = "+")
pretty_percent

# Format p_values using format = "g" and digits = 2
formatC(p_values, digits = 2, format = "g")

```


## paste()

```{r}
years <- c(2010:2013)

# Add $ to pretty_income
paste("$", pretty_income, sep = "")
 
# Add % to pretty_percent
paste(pretty_percent, "%", sep = "")

# Create vector with elements like 2010: +4.0%`
year_percent <- paste(years, ": ", pretty_percent, "%", sep = "")

# Collapse all years into single string
paste(year_percent, collapse = ", ")

```


## Ejemplo de tabla con format y paste

```{r}

# Define the names vector
income_names <- c("Year 0", "Year 1", "Year 2", "Project Lifetime")

# Create pretty_income
pretty_income <- format(income, digits = 2, big.mark = ",")

# Create dollar_income
dollar_income <- paste("$", pretty_income, sep = "")

# Create formatted_names
formatted_names <- format(income_names, justify = "right")

# Create rows
rows <- paste(formatted_names, dollar_income, sep = "   ")

# Write rows
writeLines(rows)

```


## Ejemplo 2

```{r}

toppings <- paste("adicional", 1:25)

# Randomly sample 3 toppings
my_toppings <- sample(toppings, size = 3)

# Print my_toppings
my_toppings

# Paste "and " to last element: my_toppings_and
my_toppings_and <- paste(c("","","and "), c(my_toppings), sep = "")

# Collapse with comma space: these_toppings
these_toppings <- paste(my_toppings_and, collapse = ", ")

# Add rest of sentence: my_order
my_order <- paste("I want to order a pizza with ", these_toppings, ".", sep = "")

# Order pizza with writeLines()
writeLines(my_order)

```


# Introduccion a stringr

## paste vs stringr para valores NA

```{r}

my_toppings <- c("cheese", NA, NA)
my_toppings_and <- paste(c("", "", "and "), my_toppings, sep = "")

# Print my_toppings_and
my_toppings_and

# Use str_c() instead of paste(): my_toppings_str
my_toppings_str <- str_c(c("", "", "and "), my_toppings)

# Print my_toppings_str
my_toppings_str

# paste() my_toppings_and with collapse = ", "
paste(my_toppings_and, collapse = ", ")

# str_c() my_toppings_str with collapse = ", "
str_c(my_toppings_str, collapse = ", ")

```


## str_length() vs nchar() para factores

```{r}

# Extracting vectors for boys' and girls' names
babynames_2014 <- filter(babynames, year == 2014)
boy_names <- filter(babynames_2014, sex == "M")$name
girl_names <- filter(babynames_2014, sex == "F")$name

# Take a look at a few boy_names
head(boy_names)

# Find the length of all boy_names
boy_length <- str_length(boy_names)

# Take a look at a few lengths
head(boy_length)

# Find the length of all girl_names
girl_length <- str_length(girl_names)

# Find the difference in mean length
mean(girl_length) - mean(boy_length)

# Confirm str_length() works with factors
head(str_length(factor(boy_names)))

# nchar() produce error a diferencia de str_length()
head(nchar(boy_names))

```


## Extraer subcadenas

```{r}

# Extract first letter from boy_names
boy_first_letter <- str_sub(boy_names, 1, 1)

# Tabulate occurrences of boy_first_letter
table(boy_first_letter)
  
# Extract the last letter in boy_names, then tabulate
boy_last_letter <- str_sub(boy_names, -1, -1)
table(boy_last_letter)

# Extract the first letter in girl_names, then tabulate
girl_first_letter <- str_sub(girl_names, 1, 1)
table(girl_first_letter)

# Extract the last letter in girl_names, then tabulate
girl_last_letter <- str_sub(girl_names, -1, -1)
table(girl_last_letter)

```


## Detectar coincidencias

```{r}

boy_df <- babynames %>% filter(year == 2014, sex == "M")

# Look for pattern "zz" in boy_names
contains_zz <- str_detect(boy_names, pattern = fixed("zz"))

# Examine str() of contains_zz
str(contains_zz)

# How many names contain "zz"?
sum(contains_zz)

# Which names contain "zz"?
boy_names[contains_zz]

# Which rows in boy_df have names that contain "zz"?
boy_df[contains_zz,]

```


## Subconjunto de cadenas basados en una coincidencia

```{r}

# Find boy_names that contain "zz"
str_subset(boy_names, pattern = fixed("zz"))

# Find girl_names that contain "zz"
str_subset(girl_names, pattern = fixed("zz"))

# Find girl_names that contain "U"
starts_U <- str_subset(girl_names, pattern = fixed("U"))
starts_U

# Find girl_names that contain "U" and "z"
str_subset(starts_U, pattern = fixed("z"))

```


## Contar ocurrencias

```{r}

# Count occurrences of "a" in girl_names
number_as <- str_count(girl_names, pattern = fixed("a"))

# Count occurrences of "A" in girl_names
number_As <- str_count(girl_names, pattern = fixed("A"))

# Bar plots of number_as and number_As

number_as_As <- tibble(number_as, number_As)

ggplot(number_as_As, aes((factor(number_as)))) +
  geom_bar(width = 0.1) +
  geom_text(stat = 'count', aes(label = ..count.., vjust= 0), position = position_dodge(1),   color =   "blue") +
  labs(x = "number of ocurrences of a", y = "frequency", title = "The number of words found according \n to number of occurrences of \"a\"") +
  theme_fivethirtyeight()

ggplot(number_as_As, aes((factor(number_As)))) +
  geom_bar(width = 0.1) +
  geom_text(stat = 'count', aes(label = ..count.., vjust= 0), position = position_dodge(1),   color =   "blue") +
  labs(x = "number of ocurrences of A", y = "frequency", title = "The number of words found according \n to number of occurrences of \"A\"") +
  theme_wsj() +
  theme(
  rect = element_rect(fill = "grey92"),
  legend.key = element_blank(),
  axis.ticks = element_blank(),
  panel.grid = element_blank(),
  panel.grid.major.y = element_line(color = "white", size = 0.5, linetype = "dotted"),
  axis.text = element_text(color = "grey25"),
  plot.title = element_text(face = "italic", size = 16),
  legend.position = "none"
  )

# Find total "a" + "A"
total_as <- number_as + number_As

# girl_names with more than 4 a'sstr_subset(girl_names, total_as > 4)
girl_names[total_as > 4]

```


## Convertir vector de cadena en lista o en matriz

```{r}

# Some date data
date_ranges <- c("23.01.2017 - 29.01.2017", "30.01.2017 - 06.02.2017")

# Split dates using " - "
split_dates <- str_split(date_ranges, pattern = fixed(" - "))
split_dates

# Split dates with n and simplify specified
split_dates_n <- str_split(date_ranges, pattern = fixed(" - "), n = 2, simplify = TRUE)
split_dates_n

# Subset split_dates_n into start_dates and end_dates
start_dates <- split_dates_n[,1]

# Split start_dates into day, month and year pieces
str_split(start_dates, pattern = fixed("."), n = 3, simplify = TRUE)

## --
both_names <- c("Box, George", "Cox, David")

# Split both_names into first_names and last_names
both_names_split <- str_split(both_names, pattern = fixed(", "), n = 2, simplify = TRUE)

# Get first names
first_names <- both_names_split[, 2]

# Get last names
last_names <- both_names_split[, 1]

```


## Estadisticas de texto

```{r}

# Split lines into words
words <- str_split(lines, pattern = fixed(" "))

# Number of words per line
lapply(words, length)
  
# Number of characters in each word
word_lengths <- lapply(words, str_length)
  
# Average word length per line
lapply(word_lengths, mean)

```


## Reemplazo de cadenas

```{r}

# Some IDs
ids <- c("ID#: 192", "ID#: 118", "ID#: 001")

# Replace "ID#: " with ""
id_nums <- str_replace_all(ids, pattern = "ID#: ", replacement = "")

# Turn id_nums into numbers
id_ints <- as.numeric(id_nums)

# Some (fake) phone numbers
phone_numbers <- c("510-555-0123", "541-555-0167")

# Use str_replace() to replace "-" with " "
str_replace(phone_numbers, pattern = "-", replacement = " ")

# Use str_replace_all() to replace "-" with " "
str_replace_all(phone_numbers, pattern = "-", replacement = " ")

# Turn phone numbers into the format xxx.xxx.xxxx
str_replace_all(phone_numbers, pattern = "-", replacement = ".")

```


# Coincidencia de patrones con expresiones regulares

## Coincidir con el inicio o el final de la cadena

```{r}

# Some strings to practice with
x <- c("cat", "coat", "scotland", "tic toc")

# Print END
END

# Run me
str_view(x, pattern = START %R% "c")

# Match the strings that start with "co" 
str_view(x, pattern = START %R% "co")

# Match the strings that end with "at"
str_view(x, pattern = "at" %R% END)

# Match the string that is exactly "cat"
str_view(x, pattern = START %R% "cat" %R% END)

```


## Coincidir con cualquier caracter

```{r}

# Match two characters, where the second is a "t"
str_view(x, pattern = ANY_CHAR %R% "t")

# Match a "t" followed by any character
str_view(x, pattern = "t" %R% ANY_CHAR)

# Match two characters
str_view(x, pattern = ANY_CHAR %R% ANY_CHAR)

# Match a string with exactly three characters
str_view(x, pattern = START %R% ANY_CHAR %R% ANY_CHAR %R% ANY_CHAR  %R% END)

```


## Combinando con funciones stringr

```{r}

pattern <- "q" %R% ANY_CHAR

# Find names that have the pattern
names_with_q <- str_subset(boy_names, pattern)

# How many names were there?
length(names_with_q)

# Find part of name that matches pattern
part_with_q <- str_extract(boy_names, pattern)

# Get a table of counts
table(part_with_q)

# Did any names have the pattern more than once?
count_of_q <- str_count(boy_names, pattern)

# Get a table of counts
table(count_of_q)

# Which babies got these names?
with_q <- str_detect(boy_names, pattern)

# What fraction of babies got these names?
mean(with_q)

```


## Alternancia

```{r}

# Match Jeffrey or Geoffrey
whole_names <- or("Jeffrey", "Geoffrey")
str_view(boy_names, pattern = whole_names, match = TRUE)

# Match Jeffrey or Geoffrey, another way
common_ending <- or("Je", "Geo") %R% "ffrey"
str_view(boy_names, pattern = common_ending, match = TRUE)

# Match with alternate endings
by_parts <- or("Je", "Geo") %R% "ff" %R% or("ry", "ery", "rey", "erey") 
str_view(boy_names, pattern = by_parts, match = TRUE)

# Match names that start with Cath or Kath
ckath <- START %R% or("C", "K") %R% "ath"
str_view(girl_names, pattern = ckath, match = TRUE)

```


## Clases de caracteres

```{r}

# Create character class containing vowels
vowels <- char_class("aeiouAEIOU")

# Print vowels
vowels

# See vowels in x with str_view()
str_view(x, pattern = vowels)

# See vowels in x with str_view_all()
str_view_all(x, vowels)

# Number of vowels in boy_names
num_vowels <- str_count(boy_names, vowels)

# Number of characters in boy_names
name_length <- str_length(boy_names)

# Calc mean number of vowels
mean(num_vowels)

# Calc mean fraction of vowels per name
mean(num_vowels / name_length)

```


## Repeticion

```{r}

# Vowels from last exercise
vowels <- char_class("aeiouAEIOU")

# See names with only vowels
str_view(boy_names, 
  pattern = START %R% one_or_more(vowels) %R% END, 
  match = TRUE)

# Use `negated_char_class()` for everything but vowels
not_vowels <- negated_char_class("aeiouAEIOU")

# See names with no vowels
str_view(boy_names, 
  pattern = exactly(one_or_more(not_vowels)), 
  match = TRUE)

```

## Ejemplo 1

```{r}

contact <- c("Call me at 555-555-0191", "123 Main St", "(555) 555 0191", "Phone: 555.555.0191 Mobile: 555.555.0192")

# Create a three digit pattern
three_digits <- ANY_CHAR %R% ANY_CHAR %R% ANY_CHAR

# Test it
str_view_all(contact, pattern = three_digits)

# Create a separator pattern
separator <- char_class("-.() ")

# Test it
str_view_all(contact, pattern = separator)

# Use these components
three_digits <- DGT %R% DGT %R% DGT
four_digits <- three_digits %R% DGT
separator <- char_class("-.() ")

# Create phone pattern
phone_pattern <- optional(OPEN_PAREN) %R%
  three_digits %R%
  zero_or_more(separator) %R%
  three_digits %R% 
  zero_or_more(separator) %R%
  four_digits
        
# Test it           
str_view_all(contact, pattern = phone_pattern)

# Use this pattern
three_digits <- DGT %R% DGT %R% DGT
four_digits <- three_digits %R% DGT
separator <- char_class("-.() ")
phone_pattern <- optional(OPEN_PAREN) %R% 
  three_digits %R% 
  zero_or_more(separator) %R% 
  three_digits %R% 
  zero_or_more(separator) %R%
  four_digits
  
# Extract phone numbers
str_extract(contact, phone_pattern)

# Extract ALL phone numbers
str_extract_all(contact, phone_pattern)

```

